Approximation as Differentiation

### 2025-05-07 main/formal-structures.tex
Addition of morphisms derived from biproduct structure.
Zero objects.
Biproducts and semi-additive categories.

\begin{definition}[Biproducts and Semi-additive categories]
A category with finite products and coproducts is \emph{semi-additive} if the canonical morphisms (projections
and inclusions) give an isomorphism
\[\textstyle X \coprod Y \iso X \prod Y\] that is natural in both variables. The product/coproduct is called a
\emph{biproduct}, and the biproduct structure is denoted by $(\biprod, 0)$. The unit $0$ is both terminal and
initial and is called a \emph{zero} object.
\end{definition}

@@

\noindent Here $\diag$ denotes the diagonal $\prodM{\id_X}{\id_X}$ given by the universal property of the
product and $\codiag$ denotes the codiagonal $\coprodM{\id_X}{\id_Y}$ given by the universal property of the
coproduct. The $f \oplus g$ morphism is the component-wise map $\prodM{f}{g} = \coprodM{f}{g}$.

>>>>

\begin{definition}[Biproducts and Semi-additive categories]
A category with finite products and coproducts is \emph{semi-additive} if the canonical morphisms (projections
and injections) give an isomorphism
\[\textstyle X \coprod Y \iso X \prod Y\] that is natural in both variables. The product/coproduct is called a
\emph{biproduct}, and the biproduct structure is denoted by $(\biprod, 0)$. The unit $0$ is both terminal and
initial and is called a \emph{zero} object.
\end{definition}

@@

\noindent Here $\diag$ denotes the diagonal $\prodM{\id_X}{\id_X}$ given by the universal property of the
product and $\codiag$ denotes the codiagonal $\coprodM{\id_X}{\id_Y}$ given by the universal property of the
coproduct. The $f \oplus g$ morphism is the component-wise map $\prodM{f \comp \pi_1}{g \comp \pi_2} =
\coprodM{\inj_1 \comp f}{\inj_2 \comp g}$. Similarly we can exhibit a \emph{zero} morphism $0_{X,Y}$ by
composing the unique maps in and out of the zero object:

\begin{center}
\begin{tikzcd}
   X \arrow[r, "!_X"] & 0 \arrow[r, "!^Y"] & Y
\end{tikzcd}
\end{center}

It is easy to verify that $+$ is associative and commutative and that $f + 0 = f = 0 + f$, and so every
hom-object $\cat{C}(X,Y)$ is an object in $\CMon$. Moreover composition is \emph{bilinear}, i.e.~given by a
family of morphisms $\Hom{\cat{C}}{Y}{Z} \tensor \Hom{\cat{C}}{X}{Y} \to \Hom{\cat{C}}{X}{Z}$ in $\CMon$ that
preserve the additive structure in $\Hom{\cat{C}}{Y}{Z}$ and $\Hom{\cat{C}}{X}{Y}$ separately:

\begin{salign*}
f \comp \zero = f = \zero \comp f
\end{salign*}
\begin{salign*}
(f + g) \comp h &= (f \comp h) + (g \comp h) \\
h \comp (f + g) &= (h \comp f) + (h \comp g)
\end{salign*}
